Non-word repetition in Yélî Dnye

### 2021-10-18 manuscript.Rmd
update casillas2020early to casillas2021early

Our knowledge of Yélî language development is growing [e.g., @brown2011cultural; @brown2014interactional; @brownIPchildrearing; @casillas2020early; @liszkowski2012prelinguistic], but research into Yélî phonological development has only just begun. For example, Peute and Casillas [-@peuteIPconsonants] find that Yélî Dnye-learning children's early spontaneous consonant productions appear to exclusively feature simplex and typologically frequent phones. Other ongoing work on Yélî Dnye includes experiment-based infant phoneme discrimination data and errors made in elicited and spontaneous speech from young children, but these data are neither finalized nor yet externally reviewed [see @hellwigPAlanguage for more information]. These data will help better inform our current analyses based on NWR in the future (e.g., regarding common sound substitutions) but are not critical for testing our current question about the general correlation between cross-linguistic phone frequency and NWR performance.

@@

Most Yélî parents are swidden horticulturalists, who occasionally fish. Within a group of households, it is often the case that older adolescents and adults spend their day tending to their farm plots (which may not be nearby), bringing up water from the river, washing clothes, preparing food, and engaging in other such activities. Starting around age two years, children more often spend large swaths of their day playing, swimming, and foraging for fruit, nuts, and shellfish in large (~10 members) independent and mixed-age child play groups [@brownIPchildrearing; @casillas2020early]. Formal education is a priority for Yélî families, and many young parents have themselves pursued additional education beyond what is locally available [@casillas2020early]. Local schools are well out of walking distance for many children (i.e., more than 1 hour on foot or by canoe each day), so it is very common for households situated close to a school to host their school-aged relatives during the weekdays for long segments of the school year. Children start school often at around age seven, although the precise age depends on the child's readiness, as judged by their teacher.

@@

<!-- ac2mc shouldn't we totally rewrite this? // actually, this is all factually correct-->
<!-- mc2ac: while factually correct i thought it might not be a great representation of where we stand now so i made some small edits in the final sentences-->
We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors of that paper interpret as consistent with long-term effects of low levels of infant-directed speech [@cristia2020infant]. However, @cristia2020infant also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3;0 is comparably infrequent in this community [in fact it may be infrequent in many settings, including urban ones @bunceURcrosscultural], and additionally shares other societal characteristics wih the Tsimane' [e.g., is rural and relies on farming, children grow up in wide familial networks, etc; @casillas2020early]. Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour [@scaff2021daylong], and Yélî children about 3.6 minutes per hour [@casillas2020early]. Thus, if these input quantities in early childhood relate to lower NWR scores later in life, we should observe similarly low NWR scores here as in @cristia2020infant.

@@

Additionally, we estimated the usage frequency of the phones present in the target items in a corpus of child-centered recordings [@casillas2020early]. That corpus was constituted by sampling from audio-recordings (7–9 hours long), collected as 10 children aged between 1 month and 3 years went about their day.  The researchers selected 9 2.5-minute clips randomly and 11 1- or 5-minute clips by hand (selected to represent peak turn-taking and child vocal activity). These clips were segmented and transcribed by the lead researcher and a highly knowledgeable local assistant, who speaks Yélî Dnye natively, has ample experience in this kind of research, and often knew all the recorded people personally. For more details, please refer to @casillas2020early.

>>>>

Our knowledge of Yélî language development is growing [e.g., @brown2011cultural; @brown2014interactional; @brownIPchildrearing; @casillas2021early; @liszkowski2012prelinguistic], but research into Yélî phonological development has only just begun. For example, Peute and Casillas [-@peuteIPconsonants] find that Yélî Dnye-learning children's early spontaneous consonant productions appear to exclusively feature simplex and typologically frequent phones. Other ongoing work on Yélî Dnye includes experiment-based infant phoneme discrimination data and errors made in elicited and spontaneous speech from young children, but these data are neither finalized nor yet externally reviewed [see @hellwigPAlanguage for more information]. These data will help better inform our current analyses based on NWR in the future (e.g., regarding common sound substitutions) but are not critical for testing our current question about the general correlation between cross-linguistic phone frequency and NWR performance.

@@

Most Yélî parents are swidden horticulturalists, who occasionally fish. Within a group of households, it is often the case that older adolescents and adults spend their day tending to their farm plots (which may not be nearby), bringing up water from the river, washing clothes, preparing food, and engaging in other such activities. Starting around age two years, children more often spend large swaths of their day playing, swimming, and foraging for fruit, nuts, and shellfish in large (~10 members) independent and mixed-age child play groups [@brownIPchildrearing; @casillas2021early]. Formal education is a priority for Yélî families, and many young parents have themselves pursued additional education beyond what is locally available [@casillas2021early]. Local schools are well out of walking distance for many children (i.e., more than 1 hour on foot or by canoe each day), so it is very common for households situated close to a school to host their school-aged relatives during the weekdays for long segments of the school year. Children start school often at around age seven, although the precise age depends on the child's readiness, as judged by their teacher.

@@

<!-- ac2mc shouldn't we totally rewrite this? // actually, this is all factually correct-->
<!-- mc2ac: while factually correct i thought it might not be a great representation of where we stand now so i made some small edits in the final sentences-->
We add some observations that will help us integrate this study into the broader investigation of NWR across cultures. As mentioned previously, there is one report of relatively low NWR scores among the Tsimane', which the authors of that paper interpret as consistent with long-term effects of low levels of infant-directed speech [@cristia2020infant]. However, @cristia2020infant also point out that this is based on between-paper comparisons, and thus methods and myriad other factors have not been controlled for. The Yélî community can help us gain new insights into this matter because direct speech to children under 3;0 is comparably infrequent in this community [in fact it may be infrequent in many settings, including urban ones @bunceURcrosscultural], and additionally shares other societal characteristics wih the Tsimane' [e.g., is rural and relies on farming, children grow up in wide familial networks, etc; @casillas2021early]. Although infant-directed speech has been measured in different ways among the Tsimane' and the Yélî communities, our most comparable estimates at present suggest that Tsimane' young children are spoken to about 4.2 minutes per hour [@scaff2021daylong], and Yélî children about 3.6 minutes per hour [@casillas2021early]. Thus, if these input quantities in early childhood relate to lower NWR scores later in life, we should observe similarly low NWR scores here as in @cristia2020infant.

@@

Additionally, we estimated the usage frequency of the phones present in the target items in a corpus of child-centered recordings [@casillas2021early]. That corpus was constituted by sampling from audio-recordings (7–9 hours long), collected as 10 children aged between 1 month and 3 years went about their day.  The researchers selected 9 2.5-minute clips randomly and 11 1- or 5-minute clips by hand (selected to represent peak turn-taking and child vocal activity). These clips were segmented and transcribed by the lead researcher and a highly knowledgeable local assistant, who speaks Yélî Dnye natively, has ample experience in this kind of research, and often knew all the recorded people personally. For more details, please refer to @casillas2021early.
